Lesson 4
========================================================

***

### Scatterplots and Perceived Audience Size
Notes:

***

### Scatterplots
Notes:
A few bins at around age 70, 100, 111 have an abnormal amount entries with higher friend counts. Otherwise, younger facebook users tend to have more friends.

```{r Scatterplots}
setwd('~/Dropbox/Projects/da/eda/lesson4/')
pf <- read.delim('pseudo_facebook.tsv')
library(ggthemes)
theme_set(theme_economist(12))

qplot(age, friend_count, data = pf)
```

***

#### What are some things that you notice right away?
Response:

***

### ggplot Syntax
Notes:

```{r ggplot Syntax}
library(ggplot2)
ggplot(aes(age, friend_count), data = pf) + 
  geom_point() +
  xlim(13, 90)

summary(pf$age)
```

***

### Overplotting
Notes:

```{r Overplotting}
ggplot(aes(age, friend_count), data = pf) + 
  geom_jitter(alpha = 1/20) +
  xlim(13, 90)

```

#### What do you notice in the plot?
Response:
There's a clearer concentration of users with high friend counts
at around age 20, while the spike at age 69 still persists.
***

### Coord_trans()
Notes:

```{r Coord_trans()}
?coord_trans
```

#### Look up the documentation for coord_trans() and add a layer to the plot that transforms friend_count using the square root function. Create your plot!

```{r}
ggplot(aes(age, friend_count), data = pf) + 
  geom_point(alpha=1/20, position=position_jitter(h=0)) +
  xlim(13, 90) +
  coord_trans(y='sqrt')
```

#### What do you notice?

***

### Alpha and Jitter
Notes:
Explore the relationship between friends initiated and age

```{r Alpha and Jitter}
ggplot(aes(age, friendships_initiated), data = pf) + 
  geom_point(alpha=1/20, position=position_jitter(h=0)) +
  xlim(13, 90) +
  coord_trans(y='sqrt')
```

***

### Overplotting and Domain Knowledge
Notes:

***

### Conditional Means
Notes:

```{r Conditional Means}
# install.packages('dpylr')
library('dplyr')

age_groups <- group_by(pf, age)
pf.fc_by_age <- summarize(age_groups,
          friend_count_mean = mean(friend_count),
          friend_count_median = median(friend_count),
          n = n())
pf.fc_by_age <- arrange(pf.fc_by_age)
head(pf.fc_by_age)

# Alternate method: pipe operator %>%
pf.fc_by_age <- pf %>%
  group_by(age) %>%
  summarize(friend_count_mean = mean(friend_count),
            friend_count_median = median(friend_count),
            n = n()) %>%
  arrange(age)
head(pf.fc_by_age, 20)

```

Create your plot!

```{r Conditional Means Plot}
ggplot(aes(age, friendships_initiated), data = pf) + 
  geom_point(alpha=1/20,
             position=position_jitter(h=0),
             color='#099DD9') +
  xlim(13, 90) +
  coord_trans(y='sqrt') +
  geom_line(stat='summary', fun.y = mean) +
  geom_line(stat='summary', fun.y = quantile, probs = .1,
            linetype=2, color='#CC2127') +
  geom_line(stat='summary', fun.y = quantile, probs = .5,
            color='#CC2127') +
  geom_line(stat='summary', fun.y = quantile, probs = .9,
            linetype=2, color='#CC2127')
```

***

### Overlaying Summaries with Raw Data
Notes:
The median appears to be much less than the mean, indicating a skew towards lower friend counts than the mean suggests.

```{r Overlaying Summaries with Raw Data}
ggplot(aes(age, friendships_initiated), data = pf) + 
  coord_cartesian(xlim=c(13,70), ylim=c(0,1000)) +
  geom_point(alpha=1/20,
             position=position_jitter(h=0),
             color='#099DD9') +
  geom_line(stat='summary', fun.y = mean) +
  geom_line(stat='summary', fun.y = quantile, probs = .1,
            linetype=2, color='#CC2127') +
  geom_line(stat='summary', fun.y = quantile, probs = .5,
            color='#CC2127') +
  geom_line(stat='summary', fun.y = quantile, probs = .9,
            linetype=2, color='#CC2127')

ggplot(aes(age, friend_count_mean), data=pf.fc_by_age) +
  geom_line()
```

#### What are some of your observations of the plot?
Response:

***

### Moira: Histogram Summary and Scatterplot
See the Instructor Notes of this video to download Moira's paper on perceived audience size and to see the final plot.

Notes:

***

### Correlation
Notes:

```{r Correlation}
cor(pf$age, pf$friend_count)
# cor.test(pf$age, pf$friend_count, method='pearson')
# cor.test(pf$age, pf$friend_count)
# with(pf, cor.test(age, friend_count))
```

Look up the documentation for the cor.test function.

What's the correlation between age and friend count? Round to three decimal places.
Response:
-0.027
***

### Correlation on Subsets
Notes:

```{r Correlation on Subsets}
with(subset(pf, age >= 13 & age <= 70), cor.test(age, friend_count))
```

***

### Correlation Methods
Notes:

***

## Create Scatterplots
Notes:

```{r}
ggplot(aes(www_likes_received, likes_received), data=pf) +
  geom_point() +
  xlab('Likes received through desktop web') +
  ylab('Likes received')
```

***

### Strong Correlations
Notes:

```{r Strong Correlations}
ggplot(aes(www_likes_received, likes_received), data=pf) +
  geom_point() +
  xlim(0, quantile(pf$www_likes_received, .95)) +
  ylim(0, quantile(pf$likes_received, .95)) +
  geom_smooth(method='lm', color='#CC2127') +
  xlab('Likes received through desktop web') +
  ylab('Likes received')
```

What's the correlation betwen the two variables? Include the top 5% of values for the variable in the calculation and round to 3 decimal places.

```{r Correlation Calcuation}
cor.test(pf$www_likes_received, pf$likes_received)
```

Response:
0.948
***

### Moira on Correlation
Notes:

***

### More Caution with Correlation
Notes:

```{r More Caution With Correlation}
# install.packages('alr3')
library(alr3)
data(Mitchell)
?Mitchell
```

Create your plot!

```{r Temp vs Month}
ggplot(aes(Month, Temp), data=Mitchell) + geom_point()
```

***

### Noisy Scatterplots
a. Take a guess for the correlation coefficient for the scatterplot.
0

b. What is the actual correlation of the two variables?
(Round to the thousandths place)
0.057

```{r Noisy Scatterplots}
cor.test(Mitchell$Month, Mitchell$Temp)
```

***

### Making Sense of Data
Notes:

```{r Making Sense of Data}
ggplot(aes(Month, Temp), data=Mitchell) +
  geom_point() +
  scale_x_discrete(breaks=seq(0,203,12))
```

***

### A New Perspective

What do you notice?
Response:
There seems to be a cyclical pattern corresponding to the seasons of the year.

Watch the solution video and check out the Instructor Notes!
Notes:

***

### Understanding Noise: Age to Age Months
Notes:

```{r Understanding Noise: Age to Age Months}
ggplot(aes(age, friend_count_mean), data=pf.fc_by_age) +
  geom_line()

head(pf.fc_by_age)
pf.fc_by_age[17:19,]
```

***

### Age with Months Means

```{r Age with Months Means}
pf$age_with_months <- with(pf, age + (1 - dob_month / 12))
```

Programming Assignment
```{r Programming Assignment}
pf.fc_by_age_months <- pf %>%
  group_by(age_with_months) %>%
  summarize(friend_count_mean = mean(friend_count),
            friend_count_median = median(friend_count),
            n = n()) %>%
  arrange(age_with_months)

head(pf.fc_by_age_months, 20)
```

***

### Noise in Conditional Means

```{r Noise in Conditional Means}
ggplot(aes(age_with_months, friend_count_mean),
       data=subset(pf.fc_by_age_months, age_with_months < 71)) +
  geom_line()
```

***

### Smoothing Conditional Means
Notes:

```{r Smoothing Conditional Means}
p1 <- ggplot(aes(age, friend_count_mean),
       data=subset(pf.fc_by_age, age < 71)) +
  geom_line() +
  geom_smooth()

p2 <- ggplot(aes(age_with_months, friend_count_mean),
       data=subset(pf.fc_by_age_months, age_with_months < 71)) +
  geom_line() +
  geom_smooth()

p3 <- ggplot(aes(round(age/5)*5, friend_count_mean),
       data=subset(pf.fc_by_age, age < 71)) +
  geom_line(stat='summary', fun.y=mean) +
  geom_smooth()

library(gridExtra)
grid.arrange(p2, p1, p3)
```

***

### Which Plot to Choose?
Notes:

***

### Analyzing Two Variables
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

