### Exploring Prosper Loan Data

Today we're exploring Prosper's peer-to-peer lending data
collected up until 03/11/2014. Prosper is the first
peer-to-peer lending marketplace, and fortunately for us,
they maintain a full public database of all loans where
we can analyze the performance of any subset of loans.

This data set contains 113,937 loans
with 81 variables on each loan, including:
- loan amount
- borrower rate (or interest rate)
- current loan status
- borrower income
- borrower employment status
- borrower credit history
- the latest payment information
among others.

========================================================

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
# Load all of the packages that you end up using
# in your analysis in this code chunk.

# Notice that the parameter "echo" was set to FALSE for this code chunk.
# This prevents the code from displaying in the knitted HTML output.
# You should set echo=FALSE for all code chunks in your file.

library(ggplot2)
library(ggthemes)
library(dplyr)
theme_set(theme_minimal(12))
```

Let's check out the data at a glance

```{r Load_the_Data}
# Load the Data
setwd('~/Dropbox/Projects/da/eda/prosper-loan-data/')

prosper <- read.csv('prosper-loan-data.csv')
summary(prosper)
str(prosper)

defs <- read.csv('variable-definitions.csv')
```

# Univariate Plots

Here's a few plots to start off with

```{r, Prosper Rating}
# Histogram of Loans by Propser Rating
prosper$ProsperRating <-
  factor(prosper$ProsperRating..Alpha,
         levels = c('AA', 'A', 'B', 'C', 'D', 'E', 'HR', 'NA'))
ggplot(prosper, aes(ProsperRating)) +
  geom_histogram(aes(fill = ProsperRating)) +
  ggtitle('Number of Loans by Prosper Rating') +
  xlab('Prosper Rating') +
  ylab('Number of Loans')

# Histogram of Amount Loaned by Propser Rating
ggplot(prosper, aes(ProsperRating, LoanOriginalAmount)) +
  geom_bar(stat='identity', aes(fill = ProsperRating)) +
  ggtitle('Amount Loaned by Prosper Rating') +
  xlab('Prosper Rating') +
  ylab('Amount Loaned')
```

The number of loans by Prosper rating seems to be normally distributed, with a minor positive skew towards higher rated loans.

Plotting the amount, rather than count of loans against the Prosper rating seems to emphasize this.

```{r, Investor}
# Histogram of Investors Per Loan
ggplot(prosper, aes(Investors)) +
  geom_histogram(color = 'black', fill = '#007EE5') +
  ggtitle('Number of Loans by Investor Count') +
  xlab('Number of Investors') +
  ylab('Number of Loans')

ggplot(prosper, aes(Investors)) +
  scale_x_log10() +
  geom_histogram(color = 'black', fill = '#007EE5') +
  ggtitle('Number of Loans by Investor Count') +
  xlab('Number of Investors') +
  ylab('Number of Loans')
```

The number of investors per loan seem to be heavily positive skewed. Performing a log transformation to restore symmetry results in a much more normal distribution, indicating that there's really a lognormal distribution.

```{r, Category}
# Histogram of Loan Terms by Year
categories <- c('Not Available', 'Debt Consolidation', 'Home Improvement',
                'Business', 'Personal Loan', 'Student Use', 'Auto', 'Other',
                'Baby&Adoption', 'Boat', 'Cosmetic Procedure', 'Engagement Ring',
                'Green Loans', 'Household Expenses', 'Large Purchases',
                'Medical/Dental', 'Motorcycle', 'RV', 'Taxes', 'Vacation',
                'Wedding Loans')
prosper$ListingCategory <- prosper$ListingCategory..numeric %>%
  as.factor() %>%
  factor(labels = categories)
ggplot(prosper, aes(ListingCategory)) +
  geom_histogram(aes(fill = ListingCategory)) +
  ggtitle('Number of Loans by Category') +
  xlab('Category') +
  ylab('Number of Loans')
```

We see that the majority loans are debt consolidations, with small but significant amounts of home improvement and business loans.

```{r, Year and Category}
# Formatting dates
prosper$LoanOriginationDate <- prosper$LoanOriginationDate %>%
  as.Date(format = '%Y-%m-%d')

# Number of Loans by Year, Grouped by Loan Category
prosper$LoanOriginationDate.year <-
  prosper$LoanOriginationDate %>%
  format('%Y') %>%
  strtoi()
ggplot(prosper, aes(as.factor(LoanOriginationDate.year))) +
  geom_histogram(color = 'black', aes(fill = prosper$ListingCategory)) +
  ggtitle('Number of Loans by Year') +
  xlab('Year') +
  ylab('Number of Loans')
```

Batching our categories of loans to their respective years, we see categories were only introduced in late 2007. There was a dip in loans offered in 2009 corresponding to their SEC relaunch, but there's been a growth of loan offerings ever since. The dip in 2014 reflects the fact that the data was collected up until 03/11/2014

```{r, Loan Status}
# Now Grouped by Loan Status
ggplot(prosper, aes(as.factor(LoanOriginationDate.year))) +
  geom_histogram(color = 'black', aes(fill = prosper$LoanStatus)) +
  ggtitle('Number of Loans by Year') +
  xlab('Year') +
  ylab('Number of Loans')
```

Here we're using loan status instead of category. We see that default rates were significantly reduced after their 2009 relaunch. Current loans span from years 2011-2014. Keep this histogram doesn't display the distribution of loan-terms, since a recently created 36-month loan would still be placed in the current year's batch.

```{r Month}
# Histogram Showing the Growth of Number of Loans by Month
months <- c('Jan', 'Feb', 'Mar', 'Apr', 'May', 'Jun',
            'Jul', 'Aug', 'Sep', 'Oct', 'Nov', 'Dec')
prosper$LoanOriginationDate.month <- prosper$LoanOriginationDate %>%
  format('%b')
prosper$LoanOriginationDate.month <- factor(prosper$LoanOriginationDate.month,
                                            months)
ggplot(prosper, aes(LoanOriginationDate.month)) +
  geom_histogram(color = 'black', fill = '#007EE5') +
  facet_wrap(~LoanOriginationDate.year) +
  ggtitle('Number of Loans by Month') +
  xlab('Month') +
  ylab('Number of Loans')
```

Taking a step back and visualizing the growth loan offerings month-by-month, we see once again the absence of loans during 2009, and the subsequent growth ever since.

```{r Day}
# Number of Loans by Day of the Month, shows a cyclical pattern
prosper$LoanOriginationDate.day <- prosper$LoanOriginationDate %>%
  format('%d') %>%
  strtoi()
ggplot(prosper, aes(as.factor(LoanOriginationDate.day))) +
  geom_histogram(color = 'black', fill = '#00B488') +
  scale_x_discrete(breaks = seq(1, 31, 2)) +
  facet_wrap(~LoanOriginationDate.month) +
  ggtitle('Number of Loans by Day of the Month') +
  xlab('Day of the Month') +
  ylab('Number of Loans')
```

Breaking it up further by day, lets note a cyclical pattern of loan offerings. There's a spike in offerings corresponding to end of fiscal quarters, and lack or absence of offerings during weekends and holidays.

```{r Terms and State}
# Number of Loans by State, Grouped by Term
prosper$Term.f <- as.factor(prosper$Term)
ggplot(prosper, aes(BorrowerState)) +
  geom_histogram(aes(fill = prosper$Term.f)) +
  ggtitle('Number of Loans by State') +
  xlab('State') +
  ylab('Number of Loans')
```

Recall we didn't get a chance to investigate the distribution of loans by loan terms yet. Here we perform a state-wide breakdown of loans offered by loan term. We see more loan offerings the more populous the state is. Most loan-terms are 36-month, with some 60-month, and a negligible amount of 12-month terms.

```{r Prosper Rating Boxplots}
# Boxplots of loan amounts by Prosper Rating
ggplot(subset(prosper, !is.na(ProsperRating)),
       aes(ProsperRating, LoanOriginalAmount)) +
  geom_boxplot(aes(color = ProsperRating))
```

These boxplots confirm my intuition on that those with higher Prosper ratings qualify for higher loan amounts and thus tend to borrow more. Good luck if you have anything less than a C rating.


```{r echo=FALSE, Univariate_Plots}




# Number of Loans by Loan Amount
ggplot(prosper, aes(LoanOriginalAmount)) + 
  geom_histogram()

# Number of Loans by Monthly Loan Payment
ggplot(prosper, aes(MonthlyLoanPayment)) + 
  geom_histogram(binwidth = 30)

# Number of Loans by Monthly Loan Payment
#ggplot(subset(prosper, !is.na(CreditScoreRangeLower)),
#       aes(CreditScoreRangeLower)) +
#  xlim(quantile(prosper$CreditScoreRangeLower, .01), 880) +
#  geom_histogram(binwidth = 1)

summary(prosper$CreditGrade)
ggplot(subset(prosper, !is.na(CreditGrade)), aes(CreditGrade)) + 
  geom_histogram(aes(fill = EmploymentStatus))


summary(prosper$BorrowerRate)
ggplot(prosper, aes(BorrowerRate)) + 
  geom_histogram(aes(fill = EmploymentStatus))

# LoanStatus

summary(prosper$StatedMonthlyIncome)
ggplot(prosper, aes(StatedMonthlyIncome)) + 
  xlim(0, quantile(prosper$StatedMonthlyIncome, 0.99)) +
  geom_histogram()

```

# Univariate Analysis

### What is the structure of your dataset?

### What is/are the main feature(s) of interest in your dataset?

### What other features in the dataset do you think will help support your investigation into your feature(s) of interest?

### Did you create any new variables from existing variables in the dataset?

### Of the features you investigated, were there any unusual distributions? Did you perform any operations on the data to tidy, adjust, or change the form of the data? If so, why did you do this?



# Bivariate Plots Section
```{r echo=FALSE, Bivariate_Plots}
# Amount Borrowed over Time by Term
prosper$Term.f <- as.factor(prosper$Term)
ggplot(prosper, aes(LoanOriginationDate, LoanOriginalAmount, color = Term.f)) +
  geom_point(alpha = 1/10, position = 'jitter') +
  ggtitle('Amount Borrowed Over Time') +
  xlab('Year') +
  ylab('Amount Borrowed ($USD)')
# meh


# Amount Borrowed over Time w/ Mean Trendline
prosper$age_with_months <- with(prosper, 2015 - LoanOriginationDate.year + (1 - as.numeric(LoanOriginationDate.month) / 12))
ggplot(prosper, aes(age_with_months)) +
  geom_histogram()
prosper.loan_amt_by_age_months <- prosper %>%
  group_by(age_with_months) %>%
  summarize(loan_amt_mean = mean(LoanOriginalAmount),
            loan_amt_median = median(LoanOriginalAmount),
            n = n()) %>%
  arrange(age_with_months)
ggplot(prosper.loan_amt_by_age_months, aes(age_with_months, loan_amt_mean)) +
  geom_line(stat = 'summary', fun.y = mean, linetype = 2)

ggplot(prosper, aes(age_with_months, LoanOriginalAmount)) +
  geom_point(alpha = 1/50, color = '#007EE5', position = 'jitter') +
  geom_line(stat = 'summary', fun.y = mean) +
  geom_line(stat = 'summary', fun.y = quantile, probs = .1,
            linetype = 2, color = '#CC2127') +
  geom_line(stat = 'summary', fun.y = quantile, probs = .5,
            color = '#CC2127') +
  geom_line(stat = 'summary', fun.y = quantile, probs = .9,
            linetype = 2, color = '#CC2127') +
  ggtitle('Amount Borrowed Over Time') +
  xlab('Year') +
  ylab('Amount Borrowed ($USD)')


# Loan amount vs number of investors
ggplot(prosper, aes(LoanOriginalAmount, Investors)) +
  geom_point(alpha = 1/50,
             aes(color = prosper$Term.f),
             position = 'jitter')
# meh
```

# Bivariate Analysis

### Talk about some of the relationships you observed in this part of the investigation. How did the feature(s) of interest vary with other features in the dataset?

### Did you observe any interesting relationships between the other features (not the main feature(s) of interest)?

### What was the strongest relationship you found?




# Multivariate Plots Section

```{r echo=FALSE, Multivariate_Plots}

```

# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the investigation. Were there features that strengthened each other in terms of looking at your feature(s) of interest?

### Were there any interesting or surprising interactions between features?

### OPTIONAL: Did you create any models with your dataset? Discuss the strengths and limitations of your model.

------

# Final Plots and Summary

### Plot One
```{r echo=FALSE, Plot_One}
prosper$CreditGrade <-
  factor(prosper$CreditGrade,
         levels = c('AA', 'A', 'B', 'C', 'D', 'E', 'HR', 'NC', ''))
ggplot(prosper, aes(StatedMonthlyIncome), binwidth = 250) +
  geom_histogram() +
  xlim(c(0,1000)) +
  facet_wrap(~CreditGrade)
```

### Description One


### Plot Two
```{r echo=FALSE, Plot_Two}

```

### Description Two


### Plot Three
```{r echo=FALSE, Plot_Three}

```

### Description Three

------

# Reflection

