### Exploring Prosper Loan Data

This data set contains 113,937 loans
with 81 variables on each loan, including:
- loan amount
- borrower rate (or interest rate)
- current loan status
- borrower income
- borrower employment status
- borrower credit history
- the latest payment information
among others.

Data collected until 03/11/2014
========================================================

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
# Load all of the packages that you end up using
# in your analysis in this code chunk.

# Notice that the parameter "echo" was set to FALSE for this code chunk.
# This prevents the code from displaying in the knitted HTML output.
# You should set echo=FALSE for all code chunks in your file.

library(ggplot2)
library(ggthemes)
library(dplyr)
theme_set(theme_minimal(12))
```

```{r echo=FALSE, Load_the_Data}
# Load the Data
setwd('~/Dropbox/Projects/da/eda/prosper-loan-data/')

prosper <- read.csv('prosper-loan-data.csv')
str(prosper)

defs <- read.csv('variable-definitions.csv')
```

# Univariate Plots Section
```{r echo=FALSE, Univariate_Plots}
# Formatting dates
prosper$LoanOriginationDate <- prosper$LoanOriginationDate %>%
  as.Date(format = '%Y-%m-%d')


# Number of Loans by Year, Grouped by Loan Status
prosper$LoanOriginationDate.year <-
  prosper$LoanOriginationDate %>% format('%Y')
ggplot(prosper, aes(LoanOriginationDate.year)) +
  geom_histogram(aes(fill = prosper$LoanStatus)) +
  ggtitle('Number of Loans by Year') +
  xlab('Year') +
  ylab('Number of Loans')


# Number of Loans by Month
months <- c('Jan', 'Feb', 'Mar', 'Apr', 'May', 'Jun',
            'Jul', 'Aug', 'Sep', 'Oct', 'Nov', 'Dec')
prosper$LoanOriginationDate.month <- prosper$LoanOriginationDate %>%
  format('%b')
prosper$LoanOriginationDate.month <- factor(prosper$LoanOriginationDate.month,
                                            months)
ggplot(prosper, aes(LoanOriginationDate.month)) +
  geom_histogram(fill = '#007EE5') +
  facet_wrap(~LoanOriginationDate.year) +
  ggtitle('Number of Loans by Month') +
  xlab('Month') +
  ylab('Number of Loans')


# Number of Loans by Day of the Month
prosper$LoanOriginationDate.day <- prosper$LoanOriginationDate %>%
  format('%d')
ggplot(prosper, aes(LoanOriginationDate.day)) +
  geom_histogram(fill = '#00B488') +
  scale_x_discrete(breaks = seq(1, 31, 2)) +
  facet_wrap(~LoanOriginationDate.month) +
  ggtitle('Number of Loans by Day of the Month') +
  xlab('Day of the Month') +
  ylab('Number of Loans')

# Number of Loans by State, Grouped by Term
prosper$Term.f <- as.factor(prosper$Term)
ggplot(prosper, aes(BorrowerState)) +
  geom_histogram(aes(fill = prosper$Term.f)) +
  ggtitle('Number of Loans by State') +
  xlab('State') +
  ylab('Number of Loans')




summary(prosper$CreditGrade)
ggplot(subset(prosper, !is.na(CreditGrade)), aes(CreditGrade)) + 
  geom_histogram()

summary(prosper$LoanOriginalAmount)
ggplot(prosper, aes(LoanOriginalAmount)) + 
  geom_histogram()

summary(prosper$BorrowerRate)
ggplot(prosper, aes(BorrowerRate)) + 
  geom_histogram() +
  facet_wrap(~EmploymentStatus)

# LoanStatus

summary(prosper$StatedMonthlyIncome)
ggplot(prosper, aes(StatedMonthlyIncome)) + 
  xlim(0, quantile(prosper$StatedMonthlyIncome, 0.99)) +
  geom_histogram()

# EmploymentStatus

summary(prosper$MonthlyLoanPayment)
ggplot(prosper, aes(MonthlyLoanPayment)) + 
  xlim(0, quantile(prosper$MonthlyLoanPayment, 0.99)) +
  geom_histogram()

```

# Univariate Analysis

### What is the structure of your dataset?

### What is/are the main feature(s) of interest in your dataset?

### What other features in the dataset do you think will help support your investigation into your feature(s) of interest?

### Did you create any new variables from existing variables in the dataset?

### Of the features you investigated, were there any unusual distributions? Did you perform any operations on the data to tidy, adjust, or change the form of the data? If so, why did you do this?



# Bivariate Plots Section
```{r echo=FALSE, Bivariate_Plots}
# Borrowed Amount by Month, Faceted by Year
ggplot(prosper, aes(LoanOriginationDate.month, LoanOriginalAmount)) +
  geom_bar(stat = 'identity', fill = '#007EE5') +
  facet_wrap(~LoanOriginationDate.year) +
  ggtitle('Amount Borrowed by Month') +
  xlab('Month') +
  ylab('Amount Borrowed ($USD)')


# Amount Borrowed over Time by Term
prosper$Term.f <- as.factor(prosper$Term)
ggplot(prosper, aes(LoanOriginationDate, LoanOriginalAmount, color = Term.f)) +
  geom_point(alpha = 1/50, position = 'jitter') +
  ggtitle('Amount Borrowed Over Time') +
  xlab('Year') +
  ylab('Amount Borrowed ($USD)')


# Amount Borrowed over Time w/ Mean Trendline
prosper$age_with_months <- with(prosper, 2015 - LoanOriginationDate.year + (1 - LoanOriginationDate.year / 12))
prosper.loan_amt_by_age_months <- prosper %>%
  group_by(age_with_months) %>%
  summarize(friend_count_mean = mean(friend_count),
            friend_count_median = median(friend_count),
            n = n()) %>%
  arrange(age_with_months)

ggplot(prosper, aes(LoanOriginationDate, LoanOriginalAmount)) +
  geom_point(alpha = 1/50, position = 'jitter') +
  geom_line(stat = 'summary', fun.y = mean, color = '#00B488') +
  ggtitle('Amount Borrowed Over Time') +
  xlab('Year') +
  ylab('Amount Borrowed ($USD)')

ggplot(prosper, aes(LoanOriginalAmount, Investors)) +
  geom_point(alpha = 1/50,
             aes(color = prosper$Term.f),
             position = 'jitter')
```

# Bivariate Analysis

### Talk about some of the relationships you observed in this part of the investigation. How did the feature(s) of interest vary with other features in the dataset?

### Did you observe any interesting relationships between the other features (not the main feature(s) of interest)?

### What was the strongest relationship you found?




# Multivariate Plots Section

```{r echo=FALSE, Multivariate_Plots}

```

# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the investigation. Were there features that strengthened each other in terms of looking at your feature(s) of interest?

### Were there any interesting or surprising interactions between features?

### OPTIONAL: Did you create any models with your dataset? Discuss the strengths and limitations of your model.

------

# Final Plots and Summary

### Plot One
```{r echo=FALSE, Plot_One}

```

### Description One


### Plot Two
```{r echo=FALSE, Plot_Two}

```

### Description Two


### Plot Three
```{r echo=FALSE, Plot_Three}

```

### Description Three

------

# Reflection

