Lesson 6
========================================================

### Welcome
Notes:

***

### Scatterplot Review

```{r Scatterplot Review}
setwd('/Users/James/Dropbox/Projects/da/eda/lesson6')
library(ggplot2)
library(ggthemes)
theme_set(theme_fivethirtyeight(12))
data(diamonds)

ggplot(diamonds, aes(carat, price)) +
  xlim(0, quantile(diamonds$carat, 0.99)) +
  ylim(0, quantile(diamonds$price, 0.99)) +
  geom_point(fill = '#CC2127', shape = 21) +
  stat_smooth(method = 'lm', color = '#007EE5') # linear trendline
```

***

### Price and Carat Relationship
Response:
Heavier diamonds seem to cost more.
It looks like a non-linear, possibly exponential growth relationship.
The dispersion seems to increase with higher weights.
***

### Frances Gerety
Notes:

#### A diamonds is


***

### The Rise of Diamonds
Notes:

***

### ggpairs Function
Notes:
The plot matrix uses the type of plot appropriate for the type of data, (qualitative/quantitative)

```{r ggpairs Function}
# install these if necessary
# install.packages('GGally')
# install.packages('scales')
# install.packages('memisc')
# install.packages('lattice')
# install.packages('MASS')
# install.packages('car')
# install.packages('reshape2')
# install.packages('dplyr')

# load the ggplot graphics package and the others
library(ggplot2)
library(GGally)
library(scales)
library(memisc)

# sample 10,000 diamonds from the data set
set.seed(20022012)
diamond_samp <- diamonds[sample(1:length(diamonds$price), 10000), ]
ggpairs(diamond_samp, params = c(shape = I('.'), outlier.shape = I('.')))
```

What are some things you notice in the ggpairs output?
Response:

***

### The Demand of Diamonds
Notes:

```{r The Demand of Diamonds}
# histogram of price
plot1 <- ggplot(diamonds, aes(price, fill = '#CC2127')) + 
  geom_histogram(binwidth = 100)
  ggtitle('Price')

# transform the price variable using log10
plot2 <- ggplot(diamonds, aes(price, fill = '#007EE5')) + 
  geom_histogram(binwidth = 0.01) +
  scale_x_log10() +
  ggtitle('Price (log10)')

library(gridExtra)
library(grid)
grid.arrange(plot1, plot2)
```

***

### Connecting Demand and Price Distributions
Notes:
The log plot emphasizes a more normal bimodal distribution, while the standard plot skews heavily towards the right.
***

### Scatterplot Transformation

```{r Scatterplot Transformation}
ggplot(diamonds, aes(carat, price)) +
  geom_point(alpha = 1/20, color = '#007EE5') +
  scale_x_continuous(trans = log10_trans()) +
  ggtitle('Price (log10) by Carat')
```


### Create a new function to transform the carat variable

```{r cuberoot transformation}
cuberoot_trans = function() trans_new('cuberoot',
                                      transform = function(x) x^(1/3),
                                      inverse = function(x) x^3)
```

#### Use the cuberoot_trans function
```{r Use cuberoot_trans}
ggplot(aes(carat, price), data = diamonds) + 
  geom_point() + 
  scale_x_continuous(trans = cuberoot_trans(), limits = c(0.2, 3),
                     breaks = c(0.2, 0.5, 1, 2, 3)) + 
  scale_y_continuous(trans = log10_trans(), limits = c(350, 15000),
                     breaks = c(350, 1000, 5000, 10000, 15000)) +
  ggtitle('Price (log10) by Cube-Root of Carat')
```

***

### Overplotting Revisited

```{r Sort and Head Tables}
head(sort(table(diamonds$carat), decreasing = T))
head(sort(table(diamonds$price), decreasing = T))
```


```{r Overplotting Revisited}
ggplot(aes(carat, price), data = diamonds) + 
  geom_point(alpha = 1/2, size = 3/4, position = 'jitter') + 
  scale_x_continuous(trans = cuberoot_trans(), limits = c(0.2, 3),
                     breaks = c(0.2, 0.5, 1, 2, 3)) + 
  scale_y_continuous(trans = log10_trans(), limits = c(350, 15000),
                     breaks = c(350, 1000, 5000, 10000, 15000)) +
  ggtitle('Price (log10) by Cube-Root of Carat')
```

***

### Other Qualitative Factors
Notes:

***

### Price vs. Carat and Clarity

Alter the code below.
```{r Price vs. Carat and Clarity}
# install and load the RColorBrewer package
# install.packages('RColorBrewer')
library(RColorBrewer)

ggplot(aes(x = carat, y = price, color = clarity), data = diamonds) + 
  geom_point(alpha = 0.5, size = 1, position = 'jitter') +
  scale_color_brewer(type = 'div',
    guide = guide_legend(title = 'Clarity', reverse = T,
    override.aes = list(alpha = 1, size = 2))) +  
  scale_x_continuous(trans = cuberoot_trans(), limits = c(0.2, 3),
    breaks = c(0.2, 0.5, 1, 2, 3)) + 
  scale_y_continuous(trans = log10_trans(), limits = c(350, 15000),
    breaks = c(350, 1000, 5000, 10000, 15000)) +
  ggtitle('Price (log10) by Cube-Root of Carat and Clarity')
```

***

### Clarity and Price
Response:
Yes, controlling for carat, clearer diamonds exhibit higher prices.
***

### Price vs. Carat and Cut

Alter the code below.
```{r Price vs. Carat and Cut}
ggplot(aes(x = carat, y = price, color = cut), data = diamonds) + 
  geom_point(alpha = 0.5, size = 1, position = 'jitter') +
  scale_color_brewer(type = 'div',
                     guide = guide_legend(title = 'Cut', reverse = T,
                                          override.aes = list(alpha = 1, size = 2))) +  
  scale_x_continuous(trans = cuberoot_trans(), limits = c(0.2, 3),
                     breaks = c(0.2, 0.5, 1, 2, 3)) + 
  scale_y_continuous(trans = log10_trans(), limits = c(350, 15000),
                     breaks = c(350, 1000, 5000, 10000, 15000)) +
  ggtitle('Price (log10) by Cube-Root of Carat and Cut')
```

***

### Cut and Price
Response:
Somewhat. Better cuts exhibit more variance in price,
but most diamonds are close to ideal anyways.
***

### Price vs. Carat and Color

Alter the code below.
```{r Price vs. Carat and Color}
ggplot(aes(x = carat, y = price, color = color), data = diamonds) + 
  geom_point(alpha = 0.5, size = 1, position = 'jitter') +
  scale_color_brewer(type = 'div',
                     guide = guide_legend(title = 'Color',
                                          override.aes = list(alpha = 1, size = 2))) +  
  scale_x_continuous(trans = cuberoot_trans(), limits = c(0.2, 3),
                     breaks = c(0.2, 0.5, 1, 2, 3)) + 
  scale_y_continuous(trans = log10_trans(), limits = c(350, 15000),
                     breaks = c(350, 1000, 5000, 10000, 15000)) +
  ggtitle('Price (log10) by Cube-Root of Carat and Color')
```

***

### Color and Price
Response:
Most definitely, controlling for carat, better colors exhibit higher prices.
***

### Linear Models in R
Notes:
lm(y ~ x), where y = outcome variable, x = explanatory variable

Response:
lm( log(price) ~ carat^(1/3) )
***

### Building the Linear Model
Notes:

```{r Building the Linear Model}
m1 <- lm(I(log(price)) ~ I(carat^(1/3)), data = diamonds)
m2 <- update(m1, ~ . + carat)
m3 <- update(m2, ~ . + cut)
m4 <- update(m3, ~ . + color)
m5 <- update(m4, ~ . + clarity)
mtable(m1, m2, m3, m4, m5)
```

Notice how adding cut to our model does not help explain much of the variance
in the price of diamonds. This fits with out exploration earlier.

***

### Model Problems
Video Notes:

Research:
(Take some time to come up with 2-4 problems for the model)
(You should 10-20 min on this)

Response:
The timing that the prices were collecte werent held constant,
so it may be subject to inflation and market changes
***

### A Bigger, Better Data Set
Notes:

```{r A Bigger, Better Data Set}
#install.packages('bitops')
#install.packages('RCurl')
#library('bitops')
#library('RCurl')

#diamondsurl = getBinaryURL("https://raw.github.com/solomonm/diamonds-data/master/BigDiamonds.Rda")
#load(rawConnection(diamondsurl))

load("BigDiamonds.rda")
```

The code used to obtain the data is available here:
https://github.com/solomonm/diamonds-data

## Building a Model Using the Big Diamonds Data Set
Notes:

```{r Building a Model Using the Big Diamonds Data Set}
# sampling the big diamonds dataset
diamondbig_samp <- diamondsbig[sample(1:length(diamondsbig$price), 10000), ]

# 5 models
m1 <- lm(I(log(price)) ~ I(carat^(1/3)),
         data = subset(diamondbig_samp, price < 10000 & cert == 'GIA'))
m2 <- update(m1, ~ . + carat)
m3 <- update(m2, ~ . + cut)
m4 <- update(m3, ~ . + color)
m5 <- update(m4, ~ . + clarity)

mtable(m1, m2, m3, m4, m5)
```


***

## Predictions

Example Diamond from BlueNile:
Round 1.00 Very Good I VS1 $5,601

```{r}
#Be sure you’ve loaded the library memisc and have m5 saved as an object in your workspace.
thisDiamond = data.frame(carat = 1.00, cut = "V.Good",
                         color = "I", clarity="VS1")
modelEstimate = predict(m5, newdata = thisDiamond,
                        interval="prediction", level = .95)

exp(modelEstimate)
```

Evaluate how well the model predicts the BlueNile diamond's price. Think about the fitted point estimate as well as the 95% CI.

The fitted point estimate of $5046.85 is quite close
to the actual price $5601.

The 95% CI of (3719.35, 6848.16) is quite generous,
and the actual price is within these bounds.
***

## Final Thoughts
Notes:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

