Lesson 5
========================================================

### Multivariate Data
Notes:

***

### Moira Perceived Audience Size Colored by Age
Notes:

***

### Third Qualitative Variable
Notes:

```{r Third Qualitative Variable}
setwd('~/Dropbox/Projects/da/eda/lesson5')
library(ggplot2)
library(ggthemes)
theme_set(theme_fivethirtyeight(12))

pf <- read.delim('pseudo_facebook.tsv')
# Histograms of friend counts by gender
ggplot(aes(x = friend_count),
       data = subset(pf, !is.na(gender)), binwidth=25) +
  geom_histogram() +
  scale_x_continuous(limits=c(0, 1000),
                     breaks=seq(0, 1000, 50)) +
  facet_wrap(~gender)
  
# Boxplots of friend counts by gender with mean friend count marked
ggplot(aes(x = gender, y = age),
       data = subset(pf, !is.na(gender))) +
  geom_boxplot() +
  stat_summary(fun.y=mean, geom='point', shape=4)

# Median friend counts vs age, by gender
ggplot(aes(x = age, y = friend_count),
       data = subset(pf, !is.na(gender))) +
  geom_line(aes(color=gender), stat='summary', fun.y=median)

# Create dataframe of mean & median friend counts
# by age and gender
library(dplyr)
pf.fc_by_age_gender <- pf %>%
  filter(!is.na(gender)) %>% 
  group_by(age, gender) %>%
  summarize(mean_friend_count=mean(friend_count),
            median_friend_count=median(friend_count),
            n=n())

head(pf.fc_by_age_gender)
```

***

### Plotting Conditional Summaries
Notes:

```{r Plotting Conditional Summaries}
ggplot(aes(x = age, y = median_friend_count),
       data = pf.fc_by_age_gender) +
  geom_line(aes(color=gender))
```

***

### Thinking in Ratios
Notes:

***

### Wide and Long Format
Notes:

***

### Reshaping Data
Notes:

```{r}
# install.packages('reshape2')
library(reshape2)

# Reshape long data to flat, spliting the gender data
# to respective female & male columns
pf.fc_by_age_gender.wide <-
  dcast(pf.fc_by_age_gender,
        age ~ gender,
        value.var = 'median_friend_count')

head(pf.fc_by_age_gender.wide)

# Alternatively, use dply and tidyr
library(tidyr)
pf.fc_by_age_gender.wide <-
  subset(pf.fc_by_age_gender[c('age', 'gender', 'median_friend_count')],
         !is.na(gender)) %>% 
    spread(gender, median_friend_count) %>% 
    mutate(ratio = female / male)

head(pf.fc_by_age_gender.wide)
```


***

### Ratio Plot
Notes:

```{r Ratio Plot}
ggplot(aes(x = age, y = female / male),
       data = pf.fc_by_age_gender.wide) +
  geom_line() +
  geom_hline(yintercept = 1, alpha = 0.5, linetype = 2)
```

***

### Third Quantitative Variable
Notes:

```{r Third Quantitative Variable}
pf$year_joined <- floor(2014 - pf$tenure / 365)
```

***

### Cut a Variable
Notes:

```{r Cut a Variable}
summary(pf$year_joined)
table(pf$year_joined)
?cut

# cutting year joined into the following bins
# (2004-2009]
# (2009-2011]
# (2011-2012]
# (2012-2014]

pf$year_joined.bucket <- cut(pf$year_joined,
                             breaks=c(2004, 2009, 2011, 2012, 2014))
```

***

### Plotting it All Together
Notes:

```{r Plotting it All Together}
table(pf$year_joined.bucket, useNA = 'ifany')

# Age vs friend count for each year joined bucket
ggplot(aes(x = age, y = friend_count),
       data = subset(pf, !is.na(year_joined.bucket))) +
  geom_line(aes(color = year_joined.bucket), stat ='summary', fun.y = median)
```

***

### Plot the Grand Mean
Notes:
Against the means of each bucket

```{r Plot the Grand Mean}
ggplot(aes(x = age, y = friend_count),
       data = subset(pf, !is.na(year_joined.bucket))) +
  geom_line(aes(color = year_joined.bucket), stat ='summary', fun.y = mean) +
  geom_line(stat='summary', fun.y = mean, linetype = 2)
```

***

### Friending Rate
Notes:

```{r Friending Rate}
with(subset(pf, tenure > 0), summary(friend_count / tenure))
```

***

### Friendships Initiated
Notes:

What is the median friend rate?

What is the maximum friend rate?

```{r Friendships Initiated}
with(subset(pf, tenure > 0), summary(friendships_initiated / tenure))

ggplot(subset(pf, tenure > 0), aes(tenure, friendships_initiated / tenure)) +
  geom_line(aes(color = year_joined.bucket), stat = 'summary', fun.y = mean)
```

***

### Bias-Variance Tradeoff Revisited
Notes:

```{r Bias-Variance Tradeoff Revisited}

ggplot(aes(x = tenure, y = friendships_initiated / tenure),
       data = subset(pf, tenure >= 1)) +
  geom_line(aes(color = year_joined.bucket),
            stat = 'summary',
            fun.y = mean)

ggplot(aes(x = 7 * round(tenure / 7), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

ggplot(aes(x = 30 * round(tenure / 30), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

ggplot(aes(x = 90 * round(tenure / 90), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

ggplot(aes(x = 7 * round(tenure / 7), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_smooth(aes(color = year_joined.bucket))
```

***

### Sean's NFL Fan Sentiment Study
Notes:

***

### Introducing the Yogurt Data Set
Notes:

***

### Histograms Revisited
Notes:

```{r Histograms Revisited}
yo <- read.csv('yogurt.csv')
str(yo)

# Change the id from an int to a factor
yo$id <- factor(yo$id)
str(yo)

ggplot(yo, aes(price)) + geom_histogram()
```

***

### Number of Purchases
Notes:

```{r Number of Purchases}
summary(yo)
length(unique(yo$price))
table(yo$price)

# Creates a new variable called all.purchases,
# which gives the total counts of yogurt for
# each observation or household.
yo <- transform(yo, all.purchases = strawberry + blueberry +
                    pina.colada + plain + mixed.berry)
```

***

### Prices over Time
Notes:
Price seems to be going up over time, with more frequent discounted prices,
perhaps due to coupons or promos

```{r Prices over Time}
qplot(x = all.purchases, data = yo, binwidth = 1,
     fill = I('#099DD9'))

ggplot(yo, aes(time, price)) +
  geom_jitter(alpha = 1/4, shape= 21, fill = I('#F79420'))
```

***

### Sampling Observations
Notes:
It was difficult to get a sense of the price trajectory over time for households with infrequent purchases, but those households with a significant sample size showed an upward trend towards more expensive yogurts purchased.

***

### Looking at Samples of Households

```{r Looking at Sample of Households}
set.seed(5194)
sample.ids <- sample(levels(yo$id), 16)

ggplot(subset(yo, id %in% sample.ids), aes(time, price)) +
  facet_wrap(~id) +
  geom_line() +
  geom_point(aes(size=all.purchases), pch = 1)
```

***

### The Limits of Cross Sectional Data
Notes:

***

### Many Variables
Notes:

***

### Scatterplot Matrix
Notes:
more scatterplots below the diagonal. categorical variables create faceted histograms and boxplots.
***

```{r}
# install.packages("GGally")
library(GGally)

set.seed(1836)
pf_subset <- pf[, c(2:15)]
names(pf_subset)
ggpairs(pf_subset[sample.int(nrow(pf_subset), 1000),])
```

### Even More Variables
Notes:

***

```{r}
nci <- read.table("nci.tsv")
colnames(nci) <- c(1:64)
```

### Heat Maps
Notes:

***

```{r}
# melt the data in long format
library(reshape2)
nci.long.samp <- melt(as.matrix(nci[1:200,]))
names(nci.long.samp) <- c("gene", "case", "value")
head(nci.long.samp)

# make the heat map
ggplot(aes(y = gene, x = case, fill = value),
  data = nci.long.samp) +
  geom_tile() +
  scale_fill_gradientn(colours = colorRampPalette(c("blue", "red"))(100))
```


***

### Analyzing Three of More Variables
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

